Formalizing Stack Safety as a Security Property

### 2022-12-04 submission5/noexceptions.tex
Fixed tag text size

@@

Our concurrency model is fairly
simplistic, assuming a fixed number of threads each with its own dedicated processor.
\apt{Is concurrency still alive?}
We model memory-safe stack objects, but not a heap. Regions outside of
stacks can be used however the compiler likes, including as a heap, but no protection is
built in and our properties assume that if a pointer to a stack object is stored there,
it is permanently compromised.

@@

\newcommand{\freebox}{\includegraphics[height=10px]{free.png}}
\newcommand{\pubbox}{\includegraphics[height=10px]{public.png}}
\newcommand{\objbox}{\includegraphics[height=10px]{object.png}}
\newcommand{\sealbox}{\includegraphics[height=10px]{sealed.png}}

@@

To characterize this situation in a property, we define a state element to be \emph{irrelevant}
if its value can be replaced with any other value without changing the observable behavior
of the machine. More precisely, for a set of elements \(\components\),
a pair of states \(\mach\) and \(\nach\) are {\em \(\components\)-variants} if they
agree on the value of every element not in \(\components\).
In the tagged environment, we additionally require that the values of
\emph{all} elements in \(\mach\) and \(\nach\) agree on their tags. 
\apt{Right?  This does not generalize so nicely, since we have not defined the
  non-payload part of values, but it will do ok here, I think.}

The elements of \(\components\) are \emph{irrelevant}
in \(\mach\) if every variant state produces an equivalent observation trace
(see Section \ref{sec:events}); all other elements are \emph{relevant}.
We characterize \emph{observational integrity} as the property that
every relevant element that is \(\sealed\) under the callee's view is restored
to its original value at the return point.

\begin{figure}
  \includegraphics[width=\columnwidth]{variants.png}
  \caption{Integrity Violation\apt{Tags are too small to read (also in next figures)}}
  \label{fig:variant}
\end{figure}

@@

To illustrate the issues, note that {\tt f} has also stores a 0 below the stack pointer.
%
\begin{figure}
  \includegraphics[width=\columnwidth]{variants3.png}
  \caption{Return-time Confidentiality Violation}
  \label{fig:variant3}
\end{figure}
%
Now consider three elements: the address \(\SP - 4\) (blue), the address \(\SP + 12\) (yellow),
and the register {\tt a0} (red). In the first case, the execution from the variant state
has changed the value at \(\SP - 4\). But both return states agree on that value. Therefore,
while the address was updated, it does not depend on a secret value.

In the second case, the return states disagree on the value of \(\SP + 12\). But in neither
case has that value changed since the original variants. So the difference is inherited from
the original variation, and does not depend on a secret value either. This matters because,
following the return, the caller is allowed to access its stack frame, so those addresses
may legitmately influence future execution.

But in the case of {\tt a0}, the value has changed during the call (in both the original
and the variant, although only one of these would be necessary), and its final value
differs between the variants.
Therefore, it must depend on a secret (in fact, the variable {\tt secret}).
Unless {\tt a0} happens to be irrelevant in the original trace, this is a violation of what
we term {\it return-time confidentiality}.

@@

Two states are variants with respect to a set of elements, \(\components\),
if they agree on the value of every element not in \(\components\). Our
notion of non-interference involves comparing the traces of such
\(\components\)-variants. We use this to define sets of irrelevant elements.

\definition Machine states \(\mach\) and \(\nach\) are {\em \(\components\)-variants},
written \(\mach \approx_\components \nach\), if, for
all \(\component \not \in \components\), \(\mach[\component] = \nach[\component]\).
\apt{Needs revision to discuss payloads/tags.}

>>>>

\usepackage{tikz}

@@

Our concurrency model is fairly
simplistic, assuming a fixed number of threads each with its own dedicated processor.
\sna{There is no concurrency in this document. Attempting to build it in from the beginning
  was rather complicated. The question is where to put that extra complexity? It's not difficult
  from a technical standpoint, subject to the one-thread-per-processor constraint.}
We model memory-safe stack objects, but not a heap. Regions outside of
stacks can be used however the compiler likes, including as a heap, but no protection is
built in and our properties assume that if a pointer to a stack object is stored there,
it is permanently compromised.

@@

\newcommand{\freebox}{\tikz \filldraw[fill=blue] (0,0) rectangle (10px,10px);}
\newcommand{\pubbox}{\tikz \filldraw[fill=lightgray] (0,0) rectangle (10px,10px);}
\newcommand{\objbox}{\tikz \filldraw[fill=yellow] (0,0) rectangle (10px,10px);}
\newcommand{\sealbox}{\tikz \filldraw[fill=red] (0,0) rectangle (10px,10px);}

@@

To characterize this situation in a property, we define a state element to be \emph{irrelevant}
if its value can be replaced with another value without changing the observable behavior
of the machine. This replacement value must respect any enforcement-relevant constraints on
values; in the tagged setting, we define two values as variants if they share the same tag
component. Then, for a set of elements \(\components\),
a pair of states \(\mach\) and \(\nach\) are {\em \(\components\)-variants} if
they agree on the value of every element not in \(\components\), and the values
in all elements are variants.

The elements of \(\components\) are \emph{irrelevant}
in \(\mach\) if every \(\components\)-variant state produces an equivalent observation trace
(see Section \ref{sec:events}); all other elements are \emph{relevant}.
We characterize \emph{observational integrity} as the property that
every relevant element that is \(\sealed\) under the callee's view is restored
to its original value at the return point.

\begin{figure}
  \includegraphics[width=\columnwidth]{variants.png}
  \caption{Integrity Violation}
  \label{fig:variant}
\end{figure}

@@

To illustrate the issues, note that {\tt f} has also stores a 0 below the stack pointer.
%
\begin{figure}
  \includegraphics[width=\columnwidth]{variants3.png}
  \caption{Return-time Confidentiality Violation}
  \label{fig:variant3}
\end{figure}
%
Now consider three elements: the address \(\SP - 4\), the address \(\SP + 12\),
and the register {\tt a0}. The execution from the right-hand state to its return
has changed the value at \(\SP - 4\), but that value matches that of the
left-hand return. Therefore, the change does not represent a leak.

The return states disagree on the value of \(\SP + 12\). But in neither
case has that value changed since the original variants. So the difference is inherited from
the original variation, and does not represent a leak either. We do not continue executing the
variant state after return, so these values will not cause the caller to behave differently.

But in the case of {\tt a0}, the value has changed during the call (in both the original
and the variant, although only one of these would be necessary), and its final value
differs between the variants.
Therefore, it must depend on a secret (in fact, the variable {\tt secret}).
Unless {\tt a0} happens to be irrelevant, this is a violation of what
we term {\it return-time confidentiality}.

@@

For a given system, assume an equivalence class \(\sim\) on values.
Two states are variants with respect to a set of elements, \(\components\),
if they agree on the value of every element not in \(\components\) and
all of their values are equivalent under \(\sim\).
Our notion of non-interference involves comparing the traces of such
\(\components\)-variants. We use this to define sets of irrelevant elements,
and secrets in general.

\definition Machine states \(\mach\) and \(\nach\) are {\em \(\components\)-variants},
written \(\mach \approx_\components \nach\), if, for
all \(\component \not \in \components\), \(\mach[\component] = \nach[\component]\)
and for all \(\component\), \(\mach[\component] \sim \nach[\component]\).
